|

### 2020-10-07 manuscript.Rmd
including feedback

abstract: "@moffatt_inner_2020 reported the results of an experiment (N = 26 in the final sample) comparing the facial electromyographic correlates of mental rumination and distraction, following an experimentally induced stressor. Based on the absence of significant difference in the perioral muscular activity between the rumination and distraction conditions, @moffatt_inner_2020 concluded that *self-reported* inner experience was unrelated to peripheral muscular activity as assessed using surface electromyography. We suggest this conclusion is at best hasty. Indeed, concluding on the absence of an effect based on an under-powered non-significant *p*-value is strongly problematic/uninformative. Moreover, the relation between self-reports and physiological measures was not directly assessed, but only indirectly inferred from differences (or absence thereof) in group averages. Given the ample inter-individual variability in these measures (as suggested by our reanalysis), we think inferring the individual-level relation between self-reports and physiological measures from group averages is inappropriate. Given these limitations, we conclude that there is limited evidence for the main conclusion put forward by @moffatt_inner_2020 and we suggest ways forward, both from a theoretical and from a methodological perspective. Complete source code, reproducible analyses, and figures are available at https://github.com/lnalborczyk/inner_experience_EMG."

keywords: NHST, Bayesian, logical fallacy, reanalysis, inner speech, rumination, electromyography
csl: csl/apa7.csl
classoption: man, donotrepeattitle
linenumbers: yes
floatsintext: yes
link-citations: true
numbersections: false

@@

# setting seed for reproducibility
# set.seed(666)

@@

NB: You can add comments using these tags

@@

Although there are debates about the exact nature of inner speech and whether it is better described as the mere evocation of abstract linguistic representations or as the simulation of executed (overt) speech, there is a considerable amount of evidence that (expanded) inner speech production is accompanied by peripheral muscular activity in the speech effectors [for an overview, see @nalborczyk_understanding_2019]...

Given the predominantly verbal nature of rumination [e.g., @ehring_repetitive_2008; @goldwin_concreteness_2012; @goldwin_concreteness_2013; @mclaughlin_effects_2007], we previously proposed to consider rumination as a form of inner speech and to study it using the methods that have been used to study other forms of inner speech, namely, by using surface electromyography and motor interference protocols [e.g., @nalborczyk_orofacial_2017; @nalborczyk_understanding_2019; @nalborczyk_dissociating_2020; @nalborczyk_articulatory_2020]. We first showed that induced rumination was accompanied by increased facial (both over a forehead and a perioral site) muscular activity in comparison to a rest period [@nalborczyk_orofacial_2017]. However, because rumination was only compared to a rest period, it remained uncertain whether this perioral activity was specifically related to (inner) speech processes. Therefore, we ran an extension of this study, in which we compared verbal to non-verbal rumination, which suggested that the facial EMG correlates we have previously identified were not specifically related to the verbal content of the ruminative thoughts [@nalborczyk_dissociating_2020]. We discussed these findings in length and proposed several theoretical interpretations that can account for these results in the discussion section of @nalborczyk_dissociating_2020 and more extensively in @nalborczyk_understanding_2019. Although these discussions were blatantly ignored by @moffatt_inner_2020, their experimental design nevertheless had the potential to inform our understanding of the involvement of the speech motor system in different varieties of inner speech as well as to clarify the relation between the peripheral correlates of inner speech and the (self-reported) subjective experience.

The main conclusion from @moffatt_inner_2020 is that inner experience between induced rumination and distraction differs "without a change in electromyographic correlates of inner speech". In other words, they suggest that the subjective experience of inner speech is unrelated (or loosely related) to the electromyographic correlates of inner speech, which are thought to be represented mostly by the EMG amplitude recorded over the OOI and OOS muscles. However, for this in-sample observation to be of interest in an out-of-sample context (i.e., to be informative for other non-observed individuals, or said otherwise, to bring information about the population), this absence of difference has to be based on sufficiently powered sample size (given the target effect size) as well as on reliable measures. This is unlikely to be the case here, for reasons that we will present and discuss in the present article. Moreover, a simple visual exploration of the data reveals important variability between individuals in the main effect of interest. That is, some participants had higher perioral (OOS and OOI) muscular activity in the rumination condition than in the distraction condition, and some other participants showed the reverse pattern. This suggests unexplored variation in the determinants of this effect (e.g., the content of the inner experience). Indeed, the relation between the inner experience and the physiological correlates of inner speech production was only inferred from group averages. However, given the important inter-individual variability, this reasoning appears highly problematic. In the following, we explore each of these limitations and suggests ways forward, both from a theoretical and from a methodological perspective.

@@

# data reshaping (from wide to long)
# NB: 26 participants in the final dataset...

@@

df2 <- df %>%
    select(ID, order = Order, RUM_FRO_MAX_ln:BAS_OOI_MAX_ln) %>%
    na.omit %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = starts_with(c("RUM", "DIS", "BAS") ),
        names_to = "condition",
        # names_prefix = "wk",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    separate(
        col = condition,
        into = c("condition", "muscle"),
        sep = "_", extra = "merge"
        ) %>%
    # extracts muscle name
    mutate(muscle = substr(muscle, 1, 3) ) %>%
    # from long to wide
    pivot_wider(names_from = muscle, values_from = value)
```

Their final sample of participants, after data exclusion, included 26 participants (data available at https://osf.io/hj7tz/). The EMG data is depicted in Figure \@ref(fig:general) by condition (where `BAS`, `DIS`, and `RUM` refer to the baseline, distraction, and rumination conditions, respectively) and by muscle (`FRO`, `OOI`, `OOS`). This figure shows that the average natural logarithm of the EMG peak amplitude recorded over the FRO was at similar levels in the baseline and distraction conditions, but was much higher in the rumination condition. However, the average natural logarithm of the EMG peak amplitude recorded over the OOI and OOS muscles was higher than baseline in both the rumination and distraction conditions, with a slight increase from distraction to rumination (both on the mean and median).

@@

# plotting the data
df2 %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = FRO:OOI,
        names_to = "muscle",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    mutate(order = factor(order) ) %>%
    # plots the data
    ggplot(aes(x = condition, y = value) ) +
    # violin plots
    geom_violin(
        scale = "count",
        fill = "black",
        alpha = 0.15,
        colour = "white",
        position = pd,
        # adjust = 0.8,
        draw_quantiles = 0.5,
        show.legend = FALSE
        ) +
    # plotting individual data points
    geom_quasirandom(alpha = 0.3) +
    # plotting means
    stat_summary(
        fun = mean,
        geom = "line", size = 1,
        # aes(group = interaction(muscle, order) ),
        aes(group = muscle),
        position = pd,
        show.legend = FALSE
        ) +
    # plotting means
    stat_summary(
        fun = "mean",
        geom = "point", shape = 16, size = 3,
        position = pd,
        show.legend = TRUE
        ) +
    # plotting confidence intervals
    stat_summary(
        fun.data = mean_cl_normal,
        geom = "errorbar", size = 1, width = 0,
        fun.args = list(mult = 1.96),
        show.legend = FALSE,
        position = pd
        ) +
    # facetting by muscle
    facet_wrap(~muscle, ncol = 3) +
    # axis labels
    labs(x = "Condition", y = "Natural logarithm of peak amplitude") +
    theme_bw(base_size = 10) +
    scale_colour_brewer(palette = "Dark2", direction = 1) +
    scale_fill_brewer(palette = "Dark2", direction = 1)
```

To model EMG peak amplitude variations in response to the rumination and distraction inductions, we fitted a Bayesian multivariate regression model with the natural logarithm of the EMG peak amplitude as an outcome and *Condition* (baseline, rumination, distraction) as a categorical predictor. Therefore, the intercept represents the estimated natural logarithm of the EMG peak amplitude in the baseline condition, and the slopes for the rumination and distraction conditions represent deviations from the baseline. These analyses were conducted using the `brms` package [@R-brms], an `R` implementation of Bayesian multilevel models that employs the probabilistic programming language `Stan` [@carpenter_stan_2017]. We ran four chains including each 10.000 iterations and a warmup of 2.000 iterations. Posterior convergence was assessed examining autocorrelation and trace plots, as well as the Gelman-Rubin statistic. Constant effects estimates were summarised via their posterior mean and 95% credible interval. We also report Bayes factors (BFs) computed using the Savage-Dickey method.^[This method consists in taking the ratio of the posterior density at the point of interest divided by the prior density at that point [@wagenmakers_bayesian_2010].] These BFs can be interpreted as updating factors, from prior knowledge (what we knew before seeing the data) to posterior knowledge (what we know after seeing the data). A summary of the estimations from this model is presented in Table \@ref(tab:summary). This analysis revealed strong evidence for the hypothesis of a higher average EMG peak amplitude in the rumination condition as compared to the baseline condition for both the FRO and OOI muscles. However, the BFs supported the null hypothesis (i.e., no difference) between the baseline and distraction conditions for the FRO and were inconclusive for both the OOI and OOS muscles.

@@

# priors for the varying-effects model
# priors_varying <- c(
#     prior(normal(0, 10), class = Intercept, resp = "FRO"),
#     prior(normal(0, 1), class = b, resp = "FRO"),
#     prior(exponential(1), class = sigma, resp = "FRO"),
#     prior(exponential(1), class = sd, resp = "FRO"),
#     prior(normal(0, 10), class = Intercept, resp = "OOS"),
#     prior(normal(0, 1), class = b, resp = "OOS"),
#     prior(exponential(1), class = sigma, resp = "OOS"),
#     prior(exponential(1), class = sd, resp = "OOS"),
#     prior(normal(0, 10), class = Intercept, resp = "OOI"),
#     prior(normal(0, 1), class = b, resp = "OOI"),
#     prior(exponential(1), class = sigma, resp = "OOI"),
#     prior(exponential(1), class = sd, resp = "OOI")
#     )

# fitting the model
# varying_effects <- brm(
#     mvbind(FRO, OOS, OOI) ~ 1 + condition + (1 | ID),
#     family = gaussian(),
#     prior = priors_varying,
#     data = df2,
#     chains = 4, cores = detectCores(),
#     warmup = 2000, iter = 10000,
#     control = list(adapt_delta = 0.95),
#     sample_prior = TRUE,
#     file = here("models/varying_effects")
#     )

@@

# fitting the model (for computing the bf on the intercept)
varying_effects <- brm(
    mvbind(FRO, OOS, OOI) ~ 0 + Intercept + condition + (0 + Intercept | ID),
    family = gaussian(),
    prior = priors_varying,
    data = df2,
    chains = 4, cores = detectCores(),
    warmup = 2000, iter = 10000,
    control = list(adapt_delta = 0.95),
    sample_prior = TRUE,
    file = here("models/varying_effects")
    )
```

@@

summary_varying_effects <-
    # retrieve estimations from bmod5
    summary(varying_effects, prob = 0.95)$fixed %>%
    # convert to dataframe
    data.frame %>%
    # sets rownames as the first column
    rownames_to_column(var = "response") %>%
    # removing ESS information
    dplyr::select(-Bulk_ESS, -Tail_ESS) %>%
    # split response name as two columns
    # separate(col = response, into = c("response", "term"), sep = "_", extra = "merge") %>%
    # rename columns
    magrittr::set_colnames(
        c("Term", "Estimate", "SE", "Lower", "Upper", "Rhat")
        ) %>%
    # compute BF for each effect
    mutate(
        BF10 = 1 / hypothesis(
          varying_effects, glue("{Term} = 0")
          )$hypothesis$Evid.Ratio
        ) %>%
    # round BFs
    # mutate(BF01 = ifelse(BF01 < .001, "<0.001", BF01) ) %>%
    mutate(
      BF10 = ifelse(
        BF10 < 0.001 | BF10 > 1e3,
        changeSciNot(abs(BF10) ),
        round(BF10, 3)
        )
      ) %>%
    # if numeric, round to 3 decimals
    mutate_if(is.numeric, round, 3)

@@

## Concluding on the null hypothesis from under-powered null-hypothesis significance tests: what could go wrong?

@@

This problem is tackled in modern usages of null-hypothesis significance tests by ensuring that the claim under scrutiny is submitted to *severe* tests [e.g., @mayo_severe_2006; @mayo_statistical_2018]. In general terms, the strong severity principle states that we have evidence for a claim to the extent that it survives a stringent scrutiny, that is, to the extent that it survives *severe* tests. More precisely, some claim (e.g., $\theta = 0$) is said to be *severely tested* if it had great chances of being falsified, was the claim false. When a statistical test is under-powered (for detecting a given effect size), the claim under scrutiny is not strongly (severely) tested, hence it not possible to obtain strong or reliable evidence for the claim (bad test, no evidence).

More formally, we can define $\text{SEV}(T, x0, H)$, the severity with which claim $H$ passes test $T$ with outcome $x0$, and $\text{SEV}(\mu > \mu_{1}) = \Pr(d(X) \leq d(x0); \mu = \mu_{1})$ the severity of the test the claim $\mu > \mu_{1}$ was submitted to [@mayo_severe_2006; @mayo_statistical_2018]...

Anticipating the legitimate critiques on the power of their study, @moffatt_inner_2020 report the results of a (possibly ran a posteriori) power analysis using the effect size reported in @nalborczyk_orofacial_2017 of $d = 0.72$. This represents a highly optimistic estimate of the substantive effect of interest (i.e., the difference in the natural logarithm of the EMG peak amplitude between the rumination and distraction conditions) as this effect represents the standardised mean difference in EMG amplitude *between a rest and a rumination periods* [@nalborczyk_orofacial_2017].

@@

In other words, the chances of detecting the effect (was the effect actually non-null) were less than obtaining a head in a coin flip.

@@

Once again, anticipating the legitimate critique that the absence of a significant difference is not *necessarily* "significant" evidence for the absence of an effect, @moffatt_inner_2020 reported the following Bayes factor (BF) analysis:

@@

```{r bf-dance, fig.cap = "Illustrating the distribution of Bayes factors (also known as the 'Dance of the Bayes factors') in favour of the alternative hypothesis for different effect sizes (N = 26). In the left panel, the effect size is fixed to d = 0 (i.e., the null hypothesis), in the middle panel, it is fixed to d = 0.36 (i.e., the supposed target effect size in Moffatt et al., 2020), and in the right panel, the effect size is fixed to d = 0.72 (i.e., the effect size reported in Nalborczyk et al., 2017)."}
# https://julianquandt.com/post/power-analysis-by-data-simulation-in-r-part-ii/
# http://daniellakens.blogspot.com/2016/07/dance-of-bayes-factors.html
# population effect size
cohensd <- 0.36

@@

# initialises an empty array to store results
results <- data.frame(
  BF10 = numeric(nsims),
  BF10_null = numeric(nsims),
  BF10_double = numeric(nsims)
  )

@@

  # produce N simulated participants
  x <- rnorm(n = n, mean = 0, sd = 1)

  # produce N simulated participants
  y <- rnorm(n = n, mean = 0 + cohensd, sd = 1)

  # produce N simulated participants
  y_null <- rnorm(n = n, mean = 0, sd = 1)

  # produce N simulated participants
  y_double <- rnorm(n = n, mean = 0 + cohensd * 2, sd = 1)

  results$BF10[i] <- ttestBF(
    x = x, y = y, paired = TRUE, rscale = "medium"
    ) %>%
    data.frame %>%
    pull(bf)

  results$BF10_null[i] <- ttestBF(
    x = x, y = y_null, paired = TRUE, rscale = "medium"
    ) %>%
    data.frame %>%
    pull(bf)

  results$BF10_double[i] <- ttestBF(
    x = x, y = y_double, paired = TRUE, rscale = "medium"
    ) %>%
    data.frame %>%
    pull(bf)

@@

results %>%
  pivot_longer(cols = BF10:BF10_double, names_to = "type") %>%
  mutate(
    type = factor(
      type,
      levels = c("BF10_null", "BF10", "BF10_double"),
      labels = c("Cohen's d = 0", "Cohen's d = 0.36", "Cohen's d = 0.72")
      )
    ) %>%
  ggplot(aes(x = log(value), fill = type) ) +
  geom_density(alpha = 0.5, show.legend = FALSE) +
  geom_vline(xintercept = 0, linetype = 2) +
  theme_bw(base_size = 12) +
  facet_wrap(~ type, scales = "free_x") +
  labs(x = "Natural logarithm of the BF10", y = "Density") +
  scale_colour_brewer(palette = "Dark2", direction = 1) +
  scale_fill_brewer(palette = "Dark2", direction = 1)

prop_d0 <- mean(results$BF10_null > 1)
prop_d0.36 <- mean(results$BF10 > 1)
prop_d0.72 <- mean(results$BF10_double > 1)
```

@@

Second, we conducted another simulation with the aim of assessing the relation between the sample size and the value of the BF. In the previous section, we fitted a multivariate Bayesian regression model with varying-intercepts by participant and weakly informative priors on the EMG data collected by @moffatt_inner_2020. Using this model, we i) generated new datasets from the posterior predictive distribution of this model and ii) we computed the BF in favour of the alternative hypothesis ($\text{BF}_{10}$) using the `BayesFactor` package [@R-BayesFactor]. We used a "medium" prior (i.e., a scale of 1) on the scale of the Cauchy prior for the alternative hypothesis. We repeated this procedure for varying sample sizes from 20 to 200 participants (by increments of 10 participants) with $1000$ simulations (i.e., 1000 simulated datasets) for each sample size.

@@

# plotting the data
df2 %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = FRO:OOI,
        names_to = "muscle",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    mutate(order = factor(order, labels = c("RUM -> DIS", "DIS -> RUM")) ) %>%
    # plots the data
    ggplot(aes(x = condition, y = value, colour = order, fill = order) ) +
    # plotting individual data points
    geom_quasirandom(alpha = 0.3) +
    # plotting means
    stat_summary(
        fun.y = mean,
        geom = "line", size = 1,
        aes(group = interaction(muscle, order) ),
        # aes(group = muscle),
        position = pd,
        show.legend = FALSE
        ) +
    # plotting means
    stat_summary(
        fun.y = "mean",
        geom = "point", shape = 16, size = 3,
        position = pd,
        show.legend = TRUE
        ) +
    # plotting confidence intervals
    stat_summary(
        fun.data = mean_cl_normal,
        geom = "errorbar", size = 1, width = 0,
        fun.args = list(mult = 1.96),
        show.legend = FALSE,
        position = pd
        ) +
    # facetting by muscle
    facet_wrap(~muscle, ncol = 3) +
    labs(x = "Condition", y = "Natural logarithm of peak amplitude") +
    theme_bw(base_size = 10) +
    scale_colour_brewer(palette = "Dark2", direction = 1) +
    scale_fill_brewer(palette = "Dark2", direction = 1) +
    theme(
      legend.title = element_blank(),
      legend.position = "top",
      legend.background = element_rect(size = 0.5, linetype = "solid", colour = "black")
    )
```

@@

Reproducible code and figures are available at https://github.com/lnalborczyk/inner_experience_EMG.

@@

Acknowledgements will be included in the final version of this manuscript.

>>>>

abstract: "@moffatt_inner_2020 reported the results of an experiment (N = 26 in the final sample) comparing the facial electromyographic correlates of mental rumination and distraction, following an experimentally induced stressor. Based on the absence of significant difference in the perioral muscular activity between the rumination and distraction conditions, @moffatt_inner_2020 concluded that *self-reported* inner experience was unrelated to peripheral muscular activity as assessed using surface electromyography. We suggest this conclusion is at best hasty. Indeed, concluding on the absence of an effect based on an under-powered non-significant *p*-value is strongly uninformative. Moreover, the relation between self-reports and physiological measures was not directly assessed, but only indirectly inferred from differences (or absence thereof) in group averages. Given the ample inter-individual variability in these measures (as suggested by our reanalysis), we think inferring the individual-level relation between self-reports and physiological measures from group averages is inappropriate. Given these limitations, we conclude that there is limited evidence for the main conclusion put forward by @moffatt_inner_2020 and we suggest ways forward, both from a theoretical and from a methodological perspective. Complete source code, reproducible analyses, and figures are available at https://osf.io/ba3gk/."

keywords: NHST, Bayesian, logical fallacy, reanalysis, inner speech, rumination, electromyography

csl: csl/apa7.csl
classoption: man, donotrepeattitle
linenumbers: yes
floatsintext: yes
link-citations: true
numbersections: false

@@

@@

@@

Given the predominantly verbal nature of rumination [e.g., @ehring_repetitive_2008; @goldwin_concreteness_2012; @goldwin_concreteness_2013; @mclaughlin_effects_2007], we previously proposed to consider rumination as a form of inner speech and to study it using the methods that have been used to study other forms of inner speech, namely, by using surface electromyography and motor interference protocols [e.g., @nalborczyk_orofacial_2017; @nalborczyk_understanding_2019; @nalborczyk_dissociating_2020; @nalborczyk_articulatory_2020]. We first showed that induced rumination was accompanied by increased facial (both over a forehead and a perioral site) muscular activity in comparison to a rest period [@nalborczyk_orofacial_2017]. However, because rumination was only compared to a rest period, it remained uncertain whether this perioral activity was specifically related to (inner) speech processes. Therefore, we ran a follow-up study comparing verbal to non-verbal rumination, which suggested that the facial EMG correlates we have previously identified were not specifically related to the verbal content of the ruminative thoughts [@nalborczyk_dissociating_2020]. We discussed these findings in length and proposed several theoretical interpretations that can account for these results in the discussion section of @nalborczyk_dissociating_2020 and more extensively in @nalborczyk_understanding_2019. Although these discussions were ignored by @moffatt_inner_2020, their experimental design nevertheless had the potential to inform our understanding of the involvement of the speech motor system in different varieties of inner speech as well as to clarify the relation between the peripheral correlates of inner speech and the (self-reported) subjective experience.

The main conclusion from @moffatt_inner_2020 is that inner experience between induced rumination and distraction differs "without a change in electromyographic correlates of inner speech". In other words, they suggest that the subjective experience of inner speech is unrelated (or loosely related) to the electromyographic correlates of inner speech, which are thought to be represented mostly by the EMG amplitude recorded over the orbicularis oris inferior and orbicularis oris superior muscles. However, for this in-sample observation to be of interest in an out-of-sample context (i.e., to be informative for other non-observed individuals, or said otherwise, to bring information about the population), this absence of difference has to be based on sufficiently powered sample size (given the target effect size) as well as on reliable measures. This is unlikely to be the case here, for reasons that we will present and discuss in the present article. Moreover, a simple visual exploration of the data reveals important variability between individuals in the main effect of interest. That is, some participants had higher perioral muscular activity in the rumination condition than in the distraction condition, and some other participants showed the reverse pattern. This suggests unexplored variation in the determinants of this effect (e.g., the content of the inner experience). Indeed, the relation between the inner experience and the physiological correlates of inner speech production was only inferred from group averages. However, given the important inter-individual variability, this reasoning appears highly problematic. In the following, we explore each of these limitations and suggests ways forward, both from a theoretical and from a methodological perspective.

@@

# data reshaping (from wide to long)
# NB: 26 participants in the final dataset

@@

df2 <- df %>%
    select(ID, order = Order, RUM_FRO_MAX_ln:BAS_OOI_MAX_ln) %>%
    na.omit %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = starts_with(c("RUM", "DIS", "BAS") ),
        names_to = "condition",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    separate(
        col = condition,
        into = c("condition", "muscle"),
        sep = "_", extra = "merge"
        ) %>%
    # extracts muscle name
    mutate(muscle = substr(muscle, 1, 3) ) %>%
    # from long to wide
    pivot_wider(names_from = muscle, values_from = value)
```

Their final sample of participants, after data exclusion, included 26 participants (data available at https://osf.io/hj7tz/). The EMG data is depicted in Figure \@ref(fig:general) by condition (where `BAS`, `DIS`, and `RUM` refer to the baseline, distraction, and rumination conditions, respectively) and by muscle (frontalis, `FRO`; orbicularis oris inferior `OOI`; and oribicularis oris superior, `OOS`). This figure shows that the average natural logarithm of the EMG peak amplitude recorded over the FRO was at similar levels in the baseline and distraction conditions, but was much higher in the rumination condition. However, the average natural logarithm of the EMG peak amplitude recorded over the OOI and OOS muscles was higher than baseline in both the rumination and distraction conditions, with a slight increase from distraction to rumination (both on the mean and median).

@@

# plotting the data
df2 %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = FRO:OOI,
        names_to = "muscle",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    mutate(order = factor(order) ) %>%
    # plots the data
    ggplot(aes(x = condition, y = value) ) +
    # violin plots
    geom_violin(
        scale = "count",
        fill = "black",
        alpha = 0.15,
        colour = "white",
        position = pd,
        draw_quantiles = 0.5,
        show.legend = FALSE
        ) +
    # plotting individual data points
    geom_quasirandom(alpha = 0.3) +
    # plotting means
    stat_summary(
        fun = mean,
        geom = "line", size = 1,
        aes(group = muscle),
        position = pd,
        show.legend = FALSE
        ) +
    # plotting means
    stat_summary(
        fun = "mean",
        geom = "point", shape = 16, size = 3,
        position = pd,
        show.legend = TRUE
        ) +
    # plotting confidence intervals
    stat_summary(
        fun.data = mean_cl_normal,
        geom = "errorbar", size = 1, width = 0,
        fun.args = list(mult = 1.96),
        show.legend = FALSE,
        position = pd
        ) +
    # facetting by muscle
    facet_wrap(~muscle, ncol = 3) +
    # axis labels
    labs(x = "Condition", y = "Natural logarithm of peak amplitude") +
    theme_bw(base_size = 10) +
    scale_colour_brewer(palette = "Dark2", direction = 1) +
    scale_fill_brewer(palette = "Dark2", direction = 1)
```

To model EMG peak amplitude variations in response to the rumination and distraction inductions, we fitted a Bayesian multivariate regression model with the natural logarithm of the EMG peak amplitude as an outcome and *Condition* (baseline, rumination, distraction) as a categorical predictor. Therefore, the intercept represents the estimated natural logarithm of the EMG peak amplitude in the baseline condition, and the slopes for the rumination and distraction conditions represent deviations from the baseline. These analyses were conducted using the `brms` package [@R-brms], an `R` implementation of Bayesian multilevel models that employs the probabilistic programming language `Stan` [@carpenter_stan_2017]. We ran four chains including each 10.000 iterations and a warmup of 2.000 iterations. Posterior convergence was assessed examining autocorrelation and trace plots, as well as the Gelman-Rubin statistic. Constant effects estimates were summarised via their posterior mean and 95% credible interval. We also report Bayes factors (BFs) computed using the Savage-Dickey method.^[This method consists in taking the ratio of the posterior density at the point of interest divided by the prior density at that point [@wagenmakers_bayesian_2010].] These BFs can be interpreted as updating factors, from prior knowledge (what we knew before seeing the data) to posterior knowledge (what we know after seeing the data). A summary of the estimations from this model is presented in Table \@ref(tab:summary). This analysis revealed strong evidence for the hypothesis of a higher average EMG peak amplitude in the rumination condition as compared to the baseline condition for both the FRO and OOI muscles (as assessed by the BFs). However, the BFs supported the null hypothesis (i.e., no difference) between the baseline and distraction conditions for the FRO and were inconclusive for both the OOI and OOS muscles.

@@

@@

# fitting the model
varying_effects <- brm(
    mvbind(FRO, OOS, OOI) ~ 0 + Intercept + condition + (0 + Intercept | ID),
    family = gaussian(),
    prior = priors_varying,
    data = df2,
    chains = 4, cores = detectCores(),
    warmup = 2000, iter = 10000,
    control = list(adapt_delta = 0.95),
    sample_prior = TRUE,
    file = here("models/varying_effects")
    )
```

@@

summary_varying_effects <-
    # retrieve estimations from bmod5
    summary(varying_effects, prob = 0.95)$fixed %>%
    # convert to dataframe
    data.frame %>%
    # sets rownames as the first column
    rownames_to_column(var = "response") %>%
    # removing ESS information
    dplyr::select(-Bulk_ESS, -Tail_ESS) %>%
    # rename columns
    magrittr::set_colnames(
        c("Term", "Estimate", "SE", "Lower", "Upper", "Rhat")
        ) %>%
    # compute BF for each effect
    mutate(
        BF10 = 1 / hypothesis(
          varying_effects, glue("{Term} = 0")
          )$hypothesis$Evid.Ratio
        ) %>%
    # round BFs
    mutate(
      BF10 = ifelse(
        BF10 < 0.001 | BF10 > 1e3,
        changeSciNot(abs(BF10) ),
        round(BF10, 3)
        )
      ) %>%
    # if numeric, round to 3 decimals
    mutate_if(is.numeric, round, 3)

@@

## Concluding on the null hypothesis from under-powered null-hypothesis significance tests: what could possibly go wrong?

@@

This problem is tackled in modern usages of null-hypothesis significance tests by ensuring that the claim under scrutiny is submitted to *severe* tests [e.g., @mayo_severe_2006; @mayo_statistical_2018]. In general terms, the strong severity principle states that we have evidence for a claim to the extent that it survives a stringent scrutiny, that is, to the extent that it survives *severe* tests. More precisely, some claim (e.g., $\theta = 0$) is said to be *severely tested* if it had great chances of being falsified, had the claim been false. When a statistical test is under-powered (for detecting a given effect size), the claim under scrutiny is not strongly (severely) tested, hence it not possible to obtain strong or reliable evidence for the claim (bad test, no evidence).

Anticipating the legitimate critiques on the power of their study, @moffatt_inner_2020 report the results of a power analysis using the effect size reported in @nalborczyk_orofacial_2017 of $d = 0.72$. This represents a highly optimistic estimate of the substantive effect of interest (i.e., the difference in the natural logarithm of the EMG peak amplitude between the rumination and distraction conditions) as this effect represents the standardised mean difference in EMG amplitude *between a rest and a rumination periods* [@nalborczyk_orofacial_2017].

@@

@@

Once again, anticipating the legitimate critique that the absence of a significant difference is not necessarily "significant" evidence for the absence of an effect, @moffatt_inner_2020 reported the following Bayes factor (BF) analysis:

@@

```{r bf-dance, fig.cap = "Illustrating the distribution of Bayes factors in favour of the alternative hypothesis for different population effect sizes (N = 26). In the left panel, the effect size is fixed to d = 0 (i.e., the null hypothesis), in the middle panel, it is fixed to d = 0.36 (i.e., the supposed target effect size in Moffatt et al., 2020), and in the right panel, the effect size is fixed to d = 0.72 (i.e., the effect size reported in Nalborczyk et al., 2017)."}
# inspired from http://daniellakens.blogspot.com/2016/07/dance-of-bayes-factors.html
# population effect size
cohensd <- 0.36

@@

# initialises an empty array to store the results
results <- data.frame(
  BF10 = numeric(nsims),
  BF10_null = numeric(nsims),
  BF10_double = numeric(nsims)
  )

@@

  # produces N simulated participants
  x <- rnorm(n = n, mean = 0, sd = 1)

  # produces N simulated participants when d = 0.36
  y <- rnorm(n = n, mean = 0 + cohensd, sd = 1)

  # produces N simulated participants when d = 0
  y_null <- rnorm(n = n, mean = 0, sd = 1)

  # produces N simulated participants when d = 0.72
  y_double <- rnorm(n = n, mean = 0 + cohensd * 2, sd = 1)

  # computes the BF10 when d = 0.36
  results$BF10[i] <- ttestBF(
    x = x, y = y, paired = TRUE, rscale = "medium"
    ) %>%
    data.frame %>%
    pull(bf)

  # computes the BF10 when d = 0
  results$BF10_null[i] <- ttestBF(
    x = x, y = y_null, paired = TRUE, rscale = "medium"
    ) %>%
    data.frame %>%
    pull(bf)

  # computes the BF10 when d = 0.72
  results$BF10_double[i] <- ttestBF(
    x = x, y = y_double, paired = TRUE, rscale = "medium"
    ) %>%
    data.frame %>%
    pull(bf)

@@

# plotting the results
results %>%
  pivot_longer(cols = BF10:BF10_double, names_to = "type") %>%
  mutate(
    type = factor(
      type,
      levels = c("BF10_null", "BF10", "BF10_double"),
      labels = c("Cohen's d = 0", "Cohen's d = 0.36", "Cohen's d = 0.72")
      )
    ) %>%
  ggplot(aes(x = log(value), fill = type) ) +
  geom_density(alpha = 0.5, show.legend = FALSE) +
  geom_vline(xintercept = 0, linetype = 2) +
  theme_bw(base_size = 12) +
  facet_wrap(~ type, scales = "free_x") +
  labs(x = "Natural logarithm of the BF10", y = "Density") +
  scale_colour_brewer(palette = "Dark2", direction = 1) +
  scale_fill_brewer(palette = "Dark2", direction = 1)

# computing the proportion of simulated BFs above 1
prop_d0 <- mean(results$BF10_null > 1)
prop_d0.36 <- mean(results$BF10 > 1)
prop_d0.72 <- mean(results$BF10_double > 1)
```

@@

Second, we conducted another simulation with the aim of assessing the relation between the sample size and the value of the BF. In the previous section, we fitted a multivariate Bayesian regression model with varying-intercepts by participant and weakly informative priors on the EMG data collected by @moffatt_inner_2020. Using this model, we i) generated new datasets from the posterior predictive distribution of this model and ii) we computed the BF in favour of the alternative hypothesis ($\text{BF}_{10}$) using the `BayesFactor` package [@R-BayesFactor]. We used a "medium" prior (i.e., $r = \sqrt{2}/2$) on the scale of the Cauchy prior for the alternative hypothesis. We repeated this procedure for varying sample sizes from 20 to 200 participants (by increments of 10 participants) with $1000$ simulations (i.e., 1000 simulated datasets) for each sample size.

@@

# plotting the data
df2 %>%
    # reshapes data from wide to long format
    pivot_longer(
        cols = FRO:OOI,
        names_to = "muscle",
        values_to = "value",
        values_drop_na = TRUE
        ) %>%
    mutate(order = factor(order, labels = c("RUM -> DIS", "DIS -> RUM")) ) %>%
    # plots the data
    ggplot(aes(x = condition, y = value, colour = order, fill = order) ) +
    # plotting individual data points
    geom_quasirandom(alpha = 0.3) +
    # plotting means
    stat_summary(
        fun.y = mean,
        geom = "line", size = 1,
        aes(group = interaction(muscle, order) ),
        position = pd,
        show.legend = FALSE
        ) +
    # plotting means
    stat_summary(
        fun.y = "mean",
        geom = "point", shape = 16, size = 3,
        position = pd,
        show.legend = TRUE
        ) +
    # plotting confidence intervals
    stat_summary(
        fun.data = mean_cl_normal,
        geom = "errorbar", size = 1, width = 0,
        fun.args = list(mult = 1.96),
        show.legend = FALSE,
        position = pd
        ) +
    # facetting by muscle
    facet_wrap(~muscle, ncol = 3) +
    labs(x = "Condition", y = "Natural logarithm of peak amplitude") +
    theme_bw(base_size = 10) +
    scale_colour_brewer(palette = "Dark2", direction = 1) +
    scale_fill_brewer(palette = "Dark2", direction = 1) +
    theme(
      legend.title = element_blank(),
      legend.position = "top",
      legend.background = element_rect(size = 0.5, linetype = "solid", colour = "black")
    )
```

@@

Reproducible code and figures are available at https://osf.io/ba3gk/.

@@

We thank Antonio Schettino for suggesting to include the "Dance of the Bayes factors" simulation and for provding helpful comments on a previous version of this manuscript.
